---
title: "LSCA - Lettuce - ARIMA - Store 4904"
author: "Agnes Martha"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(forecast)
library(tseries)
library(ggplot2)
```

#### 1. Data Importing

Here we will import the data for store 4904.

```{r}
lettuce <- read_excel("/Users/agnes/Downloads/LSCA/Individual Project/sub-recipe data.xlsx", sheet = "store 4904")
lettuce = lettuce[, c(1, 4)]
lettuce
```

#### 2. Simple Plotting

To get a clear idea on the data set, we will plot it as below.

```{r}
# Ensure the Date column is in Date format
lettuce$Date <- as.Date(lettuce$Date)

# Create the plot
ggplot(data = lettuce, aes(x = Date, y = `Total Lettuce per Day`)) + 
  geom_line() +  # Use geom_line for a line plot
  labs(x = "Date", y = "Total Lettuce per Day", title = "Lettuce Demand") +
  theme_minimal() +
  scale_x_date(date_breaks = "1 day", date_labels = "%d-%m-%Y") + # Format the x-axis for daily dates
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) # Tilt and make the date labels smaller

```
#### 3. Turning into Time Series

```{r}

lettuce$Date <- as.Date(lettuce$Date, format = "%d %B %Y")

# Ensure 'Total Lettuce per Day' is numeric
lettuce$`Total Lettuce per Day` <- as.numeric(lettuce$`Total Lettuce per Day`)

# Assuming daily data with potential weekly seasonality
ts_lettuce <- ts(lettuce$`Total Lettuce per Day`, frequency = 7, start = c(2015, as.numeric(format(min(lettuce$Date), "%j"))))

# Try STL decomposition with a specified seasonal window instead of 'periodic'
stl_fit <- stl(ts_lettuce, s.window = 7)  # Assuming a weekly seasonality, adjust as necessary

# Plot the STL decomposition
autoplot(stl_fit)

```
In this step we will identify the start and end time of the time series.

```{r}
start(ts_lettuce)
end(ts_lettuce)
```

#### 4. Training - Test - Split

```{r}
# Setting the training data
ts_lettuce.train <- window(ts_lettuce, end = c(2030, 1))

# Setting the testing data
ts_lettuce.test <- window(ts_lettuce, start = c(2030, 2))
```

#### 5. Choice of d and D 

```{r}
# Take the first order difference
ts_lettuce.diff1 <- diff(ts_lettuce.train, differences = 1)

# Plot the differenced series
autoplot(ts_lettuce.diff1)
```

#### 6. Stationary Test

```{r}
adf.test(ts_lettuce.diff1)
pp.test(ts_lettuce.diff1)
kpss.test(ts_lettuce.diff1)
ndiffs(ts_lettuce.diff1)
```

#### 7. Seasonal Stationarity

```{r}
# seasonal stationarity, s stand for seasonality
nsdiffs(ts_lettuce.train)
```

#### 8. Choice of p, q, P, and Q

```{r}
# acf plot
ggAcf(ts_lettuce.diff1)

# Assuming you have already computed the ACF for your differenced series:
acf_result <- acf(ts_lettuce.diff1, plot = FALSE)

# The acf_result object contains the autocorrelation values and confidence interval
# Extract the autocorrelation values
acf_values <- acf_result$acf

# The confidence interval is typically set at 95%, which corresponds to roughly 2 standard deviations
# from the mean in a normal distribution. We can compute the critical value for the 95% CI as:
critical_value <- qnorm((1 + 0.95) / 2) / sqrt(length(ts_lettuce.diff1))

# Now, find which lags have autocorrelation values outside this confidence interval
significant_lags <- which(abs(acf_values) > critical_value)

# Print the significant lags
print(significant_lags)

# If you need to see the values that cross the upper confidence limit
cross_upper <- which(acf_values > critical_value)

# If you need to see the values that cross the lower confidence limit
cross_lower <- which(acf_values < -critical_value)

# Print lags that cross the upper confidence limit
print(cross_upper)

# Print lags that cross the lower confidence limit
print(cross_lower)

```

```{r}
# pacf plot
ggPacf(ts_lettuce.diff1)

# Assuming you have already computed the PACF for your differenced series:
pacf_result <- pacf(ts_lettuce.diff1, plot = FALSE)

# The pacf_result object contains the partial autocorrelation values
# Extract the partial autocorrelation values
pacf_values <- pacf_result$acf  # Note: the name of the list element will still be 'acf'

# The confidence interval for PACF can also be computed similarly
critical_value_pacf <- qnorm((1 + 0.95) / 2) / sqrt(length(ts_lettuce.diff1))

# Find which lags have partial autocorrelation values outside this confidence interval
significant_lags_pacf <- which(abs(pacf_values) > critical_value_pacf)

# Print the significant lags for PACF
print(significant_lags_pacf)

# If you need to see the values that cross the upper confidence limit for PACF
cross_upper_pacf <- which(pacf_values > critical_value_pacf)

# If you need to see the values that cross the lower confidence limit for PACF
cross_lower_pacf <- which(pacf_values < -critical_value_pacf)

# Print lags that cross the upper confidence limit for PACF
print(cross_upper_pacf)

# Print lags that cross the lower confidence limit for PACF
print(cross_lower_pacf)

```
#### 9. Choosing the ARIMA Model

```{r}
# choose optimal p and q based on information criteria
auto.arima(ts_lettuce.train, d = 1, trace = TRUE) 
```

#### 10. Candidate Model

```{r}
ts_lettuce.model <- Arima(ts_lettuce.train, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 0), period = 7), include.drift = TRUE)
```

#### 11. Residual Analysis

```{r}
checkresiduals(ts_lettuce.model)
```

#### 12. Model Evaluation

Here we will produce the actual forecast for our data for the next 14 days.

```{r}
ts_lettuce.forecast <- forecast(ts_lettuce.model, h = 14)
ts_lettuce.forecast
```

#### 13. Out of Sample Performances

In this part we will measure the accuracy of our forecast by comparing our forecasted value and test data.

```{r}

accuracy(ts_lettuce.forecast, ts_lettuce.test)

```

#### 14. Re-calibrate

By including the whole data set from the model, we will run it again.

```{r}
ts_lettuce.model2 <- Arima(ts_lettuce, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 0), period = 7), include.drift = TRUE)
```

#### 15. Actual Forecast

And the final forecast.

```{r}
ts_lettuce.forecast2 <- forecast(ts_lettuce.model2, h = 14)
autoplot(ts_lettuce.forecast2)
ts_lettuce.forecast2
```
The next analysis by using Holt-Winter will be explained in another file.